Simulating High-Dimensional Multivariate Data

### 2021-01-15 050-examples.Rmd
2nd draft. revise ch050 with new brca data.

While the exact dxd Spearman correlation matrix is not strictly admissible in our scheme (seen by the non-PSD result above), the approximation is of sufficient quality for most applications with a maximum pairwise of `r max( targets - approximations)` and average error of `r mean( targets - approximations)`.

@@

```{r ch050-quadlossECHO, echo=T, eval=F, cache=FALSE}
## Simulate random vectors equal to the sample size
n <- nrow(brca)
## convert outside for faster simulation
nb_Rho_p <- bigsimr::cor_convert( rho = nb_Rho,
                                 from = corType, to = "pearson" )
## ensure PSD
nb_Rho_p <- bigsimr::cor_nearPSD( G = nb_Rho_p )
## create m random vectors and estimate correlation
simRho <- replicate(n = m, expr = {
  tmpSim <- rvec(n = n , nb_Rho_p, nb_margins,
                 type = 'pearson', ensure_PSD = FALSE, cores = cores)
  bigsimr::cor_fast(x = tmpSim, method = corType)
}, simplify = FALSE)
## find quadratic loss at each rep
quadLoss <- sapply(simRho, rags2ridges::loss, T = nb_Rho, type = "quadratic")
```

```{r ch050-quadloss, echo=F, eval=T, cache=FALSE}
quadLossRDS <- paste0( "results/brca", round(myProb*100), "quadLoss.rds" )
if ( !file.exists( quadLossRDS) )  {
    n <- nrow(brca)
    ## convert outside for faster simulation
    nb_Rho_p <- bigsimr::cor_convert( rho = nb_Rho,
                                     from = corType, to = "pearson" )
    ## ensure PSD
    nb_Rho_p <- bigsimr::cor_nearPSD( G = nb_Rho_p )
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m,
        expr = { tmpSim <- rvec(n = n , nb_Rho_p, nb_margins,
          ensure_PSD = FALSE, cores = cores);
          bigsimr::cor_fast( x = tmpSim, method = corType )} ,
     simplify = FALSE)
    ## find quadratic loss at each rep
    quadLoss <- unlist( lapply( simRho, rags2ridges::loss,
                               T = nb_Rho, type = "quadratic"))
    saveRDS(quadLoss, quadLossRDS)
}
quadLoss <- readRDS( quadLossRDS )
```

@@

```{r ch050-quadloss-summary, echo=TRUE, eval=T, cache=FALSE}
summary(quadLoss)
```

@@

<!-- 
## Monte Carlo estimation of MSE

Monte Carlo (MC) methods offer flexible and straightfoward statistical inferenec in this typically difficult analytic setting.
MC methods are routinely used many statistical inferential tasks including estimation, hypothesis testing, error rates, and empirical interval coverage rates.
For an concise introduction to these Methos, see, for example @Rizzo2007, Ch. 6.
As a simple example, let's now employ `bigsimr` to estimate the mean squared error (MSE) of the method of moments estimators we used in section X.

Specifically, 
TYPESET THE MATH HERE. REF to MoM above.

```{r ch050-mom-mse, echo=TRUE, eval=FALSE, cache=FALSE}
## system( "rm results/mseSizes.rds")
brcaSimMSEsizesRDS <- paste0( "results/brca", round(myProb*100), "simMSEsizes.rds" )
if ( !file.exists( brcaSimMSEsizesRDS) )  {
    n <- nrow(brca)
    m <- 100
    momSizes <- replicate(n = m, expr =
                                     { tmpSim <- rvec(n, nb_Rho, brcaMargins, type = corType, ensure_PSD = TRUE, cores = cores)
                                         tmpMom <- apply( unname(as.matrix(tmpSim)), 2, estimateNegBinMoM )
                                         tmpSizes <- unlist( lapply( X = tmpMom, function(x) x$size ) )
                                     }
                          )
    sqDiffSizes <- apply( momSizes, 2, function(x) { (x  - sizes)^2 } )
    mseSizes <- rowMeans( sqDiffSizes )
    saveRDS(mseSizes, brcaSimMSEsizesRDS)
}
mseSizes <- readRDS( brcaSimMSEsizesRDS )
```

Figure X.

```{r ch050-fig-mseSizes, echo = FALSE, eval=FALSE, message=FALSE, warning=FALSE, fig.width= 8, fig.align='center', fig.cap='Real data for the 3 selected high-expressing genes'}
mseSizes <- readRDS(file = brcaSimMSEsizesRDS)
qplot(mseSizes, bins = 20) + theme_bw()
```

One could then compare the MSE of the baisc MoM estimation scheme to other strategies ( such as MLE, trimmed approaches, etc.) and decide whether a more computational costly approach is warrented or if the MoM will suffice.

your comment -->

>>>>

While the exact dxd Spearman correlation matrix is not strictly admissible in our scheme (seen by the non-PSD result above), the approximation is of sufficient quality for most applications with a maximum pairwise of `r max( targets - approximates)` and average error of `r mean( targets - approximates)`.

@@

```{r ch050-quadlossECHO, echo=T, eval=F, cache=FALSE}
## Simulate random vectors equal to the sample size
N <- nrow(brca)
## create m random vectors and estimate correlation
simRho <- replicate(n = m, expr = {
  tmpSim <- rvec(n = N , R_X, nb_margins, cores = cores)
  bigsimr::cor_fast(x = tmpSim, method = type)
}, simplify = FALSE)
## find quadratic loss at each rep
quadLoss <- sapply(simRho, rags2ridges::loss, T = R_S, type = "quadratic")
```

```{r ch050-lossResults, echo=F, eval=T, cache=FALSE}
## quadLossRDS <- paste0( "results/brca", d, "_quadLoss.rds" )
frobeniusRDS <- paste0( "results/brca", d, "_frobenius.rds" )
if ( !file.exists( frobeniusRDS) )  {
    N <- nrow(brca)
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m, expr = {
        tmpSim <- rvec(n = N , R_X, nb_margins, cores = cores)
        bigsimr::cor_fast(x = tmpSim, method = type)
    }, simplify = FALSE)
    ## find quadratic loss at each rep
    ## simRho[[1]][1:5,1:5]
    ## R_S[1:5,1:5]
    ## rags2ridges::loss(simRho[[1]], R_S, type = 'frobenius')
    frobenius <- sapply(simRho, rags2ridges::loss, T = R_S, type = "frobenius")
    saveRDS(frobenius, frobeniusRDS)
}
frobenius <- readRDS( frobeniusRDS )
```

@@

```{r ch050-loss-summary, echo=TRUE, eval=T, cache=FALSE}
summary(frobenius)
```

@@
